---
title: "Week 3 - Basic Probability; Normal Distribution; Z-Scores"
author: "BL Mosby"

output: 
  html_document: default
---

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(tidyverse)
```

## Z-Scores

This week we will learn how to compute normal probabilities. First, let's learn how to compute z-scores from raw scores.
```{r}
GPA <- c(3.2, 2.5, 4.0, 2.7, 3.4)
GPA
```

### Method 1: "By hand"
We can calculate z-scores "by hand" using R to compute the mean and standard deviation.
```{r}
mean(GPA)
sd(GPA)
```

Once we have these values, we could calculate any z-score using the formula: $$z=\frac{x-\mu}{\sigma}$$ 
For example, let's find the z-score for a GPA score of 3.2.
```{r}
(3.2-3.16)/.594138
```

Or we could use the R capabilities of working with datasets to compute them all for us at once (this will calculate a z-score for every element in our vector). 
```{r}
(GPA-mean(GPA))/sd(GPA)
```
Note, it is standard practice for many programmers to use the **Console** below, instead of creating code chunks, for simple calculations like these.


### Method 2: Using R Functions
We can use the `scale` function to compute the z-scores for each value in the vector (dataset). 
```{r}
scale(GPA, T, T)

```
Notice that `scale` also prints the mean (or center) and standard deviation (here, scale) as well as the z-scores, which are in the same order as the original vector.


## Normal Probabilities
Now, let's learn how to compute normal probabilities in R **(this means we won't need to look things up manually on the standard normal table)**.

To find normal probabilities in R we use the `pnorm` function. The *arguments* for `pnorm()` are the z-score and the probability direction, i.e. or the upper `(lower.tail=FALSE)` or lower `(lower.tail=TRUE)` tail.

Let's try the examples from async.

Find p(z > 1.5)
```{r}
#Note this is for an upper tail so we will put lower.tail=FALSE
pnorm(1.5, lower.tail=FALSE)
```
*Q: How would we interpret this?* 6.68% of z-scores are above 1.5
6.68% of this population has a z-score over 1.5
the probability of a z-score greater than 1.5 in this population is 6.68%


Find p(z < 1.5)
```{r}
pnorm(1.5, lower.tail=TRUE)
```
*Interpretation: There is a 93.32% probability of finding a score less than z=1.5.*

Remember, since the normal distribution is symmetric, it is always the case that the probabilities for the upper and lower tails of the same z-score will add up to 1 or 100%. Use the Console to verify this.

Find p(Z < -1.5)
```{r}
#Note this is for a lower tail so we will put lower.tail=TRUE
pnorm(-1.5, lower.tail=TRUE)
```
*Q: What is the relationship between this probability, and the first one we computed?*


Find P(-1<z<1)
```{r}
#Using the pnorm function here requires us to do things differently than we did in async to find this probability. Think through this code and how/why it works.
pnorm(1, lower.tail=TRUE) - pnorm(-1, lower.tail=TRUE)
```

Find p(z < -1.96) OR p(z > 1.96)
```{r}
pnorm(-1.96, lower.tail=TRUE) + pnorm(1.96, lower.tail=FALSE)

#another way
2*pnorm(1.96, lower.tail=FALSE)

#another way
2*pnorm(-1.96, lower.tail=TRUE)
```


### IQ test example
IQ tests are normed (or scaled) to have a mean of 100 and standard deviation of 15. What is the probability an individual scores greater than 110? (or above 110 or more than 110 or over 110)

Method 1 to answer this question:
```{r}
#First let's compute the z-score - formula: z = (x-Mu)/sigma, using the Console below
#Now plug that value into the pnorm function, making sure to supply the correct direction of the tail.

pnorm(0.6666666667,lower.tail=FALSE)

```


Method 2 to answer this question:
```{r}
#Alternatively, we could provide two additional arguments to pnorm (the mean and standard deviation) to skip having to manual calculate the z-score. So now or pnorm arguments are the raw x value, mean, standard deviation, tail.

pnorm(110, mean=100, sd=15, lower.tail=FALSE)
```
Note that you would only include the additional arguments (mean and standard deviation) when you are plugging in the RAW (non-scaled) score and NOT the z-score. Think of this as a shortcut because we do not have to compute the z-score first.

What is the probability an individual scores less than a 125? (or below or under or beneath)
```{r}
#We will use the mean and sd arguments to complete this calculation in one step, rather than two:
pnorm(125, mean=100, sd=15, lower.tail=TRUE)

#Also: adding labels for the mean and standard deviation is optional. HOWEVER, make sure to always use the exact order of x value, mean, standard deviation, tail. You can also substitute T and F for TRUE and FALSE.
pnorm(125, 100, 15, lower.tail = T)

```

What is the probability that an individual scores between 85 and 115?
```{r}
#Remember for compound inequalities such as this, we will use the symmetry of the normal distribution and simple arithmetic to find the probability.

pnorm(115, 100, 15, lower.tail= T) - pnorm(85, 100, 15, lower.tail = T)
```
*Q: How is this calculation related to the empirical rule?*
