---
title: "Week 13 R Code": Multiple Regression
output: html_document
---

Week 13 Live session R code

This week we will learn how to conduct a multiple linear regression (MLR), using R.

Let's start by taking a look at the phd data from async
```{r setup}
#setwd(...) #insert your path here, or use the dropdown menu to select the folder

#These are special packages we'll use to run the standardized regression and obtain the partial correlation coefficients. You will only need to run this code once per R session.
install.packages("QuantPsyc")
install.packages("ppcor")

phd_mlr <- read.table("week13_phd2.txt", header=TRUE, sep="\t")
```

Recall in async we mean centered salary, let's do that first. (You may want to review the week 1 R code on how to add a variable to an existing dataset).
```{r}
#This adds a new variable 'SALARY_mc' to the data that is computed by subtracting the mean salary from each of the salary values.
phd_mlr$SALARY_mc <- phd_mlr$SALARY - mean(phd_mlr$SALARY)

#Now, we will run the MLR with Time, Salary (centered) and citations as predictors for # of pubs
phd.mlrreg <- lm(PUBS ~ TIME + SALARY_mc + CITS, data=phd_mlr)

summary(phd.mlrreg)
```

Interpretation of the intercept: The predicted number of publications is 4.55 when time equals 0, citations equal 0, and the salary was at its mean ($54,815.76).

Interpretation of the slope for citations: For each additional citation, there was a predicted increase in publications of .04, holding all other variables in the model constant.

Final conclusion for the model: A MLR revealed that time, salary, and citations significantly predicted number of publications, F(3,58)=15.46, p<.05, R2=0.42.

```{r}
confint(phd.mlrreg)
```

If we want to obtain standardized regression coefficients, then we can use the QuantPsyc package:

```{r}
library(QuantPsyc)

lm.beta(phd.mlrreg)
```

Let's write the standardized regression equation:
$$\hat{z}_{pubs} = 0.54z_{time} + 0.15z_{salary_{mc}} + 0.05z_{cits}$$

If we want to obtain the squared semi-partial and squared partial correlations, then we can use the ppcor package:
```{r}
library(ppcor)

spcor(phd_mlr[,c("PUBS","SALARY_mc","TIME","CITS")])
```
To get the squared semi-partial correlations then we just square each value:
$${sr}_{salary}^2 = (.1073541)^2 = .0115$$
Interpretation: 1.15% of the variation in publications was accounted for by salary, beyond that which was accounted for by time & citations.

$${sr}_{time}^2 = (.4286871)^2 = .1838$$
Interpretation: 18.38% of the variation in publications was accounted for by time, beyond that which was accounted for by salary & citations.

$${sr}_{cits}^2 = (.04087143)^2 = .0017$$
Interpretation: 0.17% of the variation in publications was accounted for by citations, beyond that which was accounted for by time & salary.


Likewise, we will use the ppcor package to get the partial correlations among the variables:
```{r}
pcor(phd_mlr[,c("PUBS","SALARY_mc","TIME","CITS")])
```

$${pr}_{salary}^2 = (.1425430)^2 =.0203$$
Interpretation: of the variation in publications not explained by time and citations, 2.03% was explained by salary.

$${pr}_{time}^2 = (.49852073)^2 = .2485$$
Interpretation: of the variation in publications not explained by salary and citations, 24.85% was explained by time.

$${pr}_{cits}^2 = (.05474607)^2 = .0030$$
Interpretation: of the variation in publications not explained by salary and time, 0.30% was explained by citations. 

And, in case you're interested here is the code to run the 3D scatterplot from async.
```{r}
#install.packages("plot3D") #you only need to run this line once
require(plot3D)
attach(phd_mlr)

fit <- lm(PUBS ~ TIME+SALARY)

TIME.pred <- seq(0, 20, length.out = 30)
SALARY.pred <- seq(40000, 80000, length.out = 30)
xy <- expand.grid(TIME = TIME.pred, 
                  SALARY = SALARY.pred)
PUBS.pred <- matrix (nrow = 30, ncol = 30, 
                     data = predict(fit, newdata = data.frame(xy), interval = "prediction"))

fitpoints <- predict(fit) 

scatter3D(z = PUBS, x = TIME, y = SALARY, pch = 18, cex = 2, 
          theta = 20, phi = 20, ticktype = "detailed",
          xlab = "time", ylab = "salary", zlab = "pubs", clab = "pubs", 
          surf = list(x = TIME.pred, y = SALARY.pred, z = PUBS.pred, 
                      facets = NA, fit = fitpoints),
          colkey = list(length = 0.8, width = 0.4),            
          main = "phd")
detach(phd_mlr)
```
