---
title: "LLO 8180 Week 7 Live Session"
author: "BL Mosby"
date: "6/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This week we learn how to conduct two-sample t-tests in R.

We will start with independent samples t-test.

Let's look at sample data given for example 9.2 (p. 297) in the book: the number of calories consumed in a meal between groups asked to eat slowly or fast

```{r}
#Store the following two vectors (Slow and Fast) as a data frame named "calories"
calories <- data.frame(Slow=c(700,450,850,600,450,550), Fast=c(450,800,750,700,550,650))

View(calories)
#This will make the data frame open in a new window once you run the code chunk.
```
It is usually a good idea to quickly inspect your data frame after creating/uploading to make sure it is correctly labeled.

Recall from Week 6 the function `t.test` can be used to quickly run a Student-t distribution hypothesis test on a given set of data. To compare two (independent) samples, we will list both data vectors within the function.

```{r}
#This will run an independent samples t-test comparing the means of the Slow and Fast variables. Note how we include the assumption the two groups have equal variance.
t.test(calories$Slow, calories$Fast, var.equal=TRUE)
```

If you **really** wanted to compute using the formulas:

```{r}
#First obtain the test statistic: get the mean and sd for each group
M1<-mean(calories$Slow)
M2<-mean(calories$Fast)
s1<-sd(calories$Slow)
s2<-sd(calories$Fast)

n1<-6; n2<-6
```
Next use the objects to obtain the t-value using the formula from async:

```{r}
t <- (M1-M2)/sqrt(
  (
    (
      (n1-1)*s1^2+(n2-1)*s2^2)/(n1+n2-2))*((1/n1)+(1/n2)))
t
```

Calculating the p-value:
We still have to double the probability for two-tailed tests, and we still need to be careful with the lower.tail command.

```{r}
#Remember pt gives the probability from the t-distribution
2*pt(t, df=10, lower.tail=TRUE)
```
This should match our output from `t.test`, and show why the formulas work. :)


APA conclusion: An independent-samples t-test revealed no statistically significant difference between the number of calories consumed between the fast eaters (M=650, s=130.38) and the slow eaters (M=600, s=154.92); t(10)=-0.60, p>0.05.


Now, lets run code for RELATED samples t-test-- that is where the two comparison groups are related or dependent on each other.

Suppose have a set of pre- and post-test scores for a group of students. To compare the means before and after, we would use a RELATED samples t-test.

```{r}
#Create the data frame
grades <- data.frame(Pretest=c(68,74,55,81,72,59,78,65,80), 
                   Posttest=c(78,79,68,92,72,65,74,71,90))

#Notice here we will remove the assumption of equal variances, and indicate the samples are paired (i.e., listed in order in both vectors)
t.test(grades$Posttest, grades$Pretest, paired=TRUE)
```

A slightly different way to compare- instead of a related two-sample test, we can do a one sample t-test on the *difference* in the grades for each student:
```{r}
#Create a new vector variable called grades
grades$difference <- c(grades$Posttest-grades$Pretest)

#Don't forget to inspect the data
grades
```

```{r}
t.test(grades$difference, mu=0)
```
Notice the output values for the test statistic, df, and p-value are all the same as in the two-sample test above.

```{r}
#Or finding the test statistic and p-value, using the formulas:
dBar<-mean(grades$difference)
dsd<-sd(grades$difference)

dt <- dBar/(dsd/sqrt(9))

#Find the p-value
2*pt(dt, df=8, lower.tail=FALSE)
```
Again, we end up with the same p-value as above.

Now, since this indicates a statistically significant difference between the pre- and post-tests, we can calculate the effect size (cohen's d), and find the means for an APA write-up/conclusion.

```{r}
#Cohen's d = Dbar/SDdiff
dBar/dsd
```
Effect size greater than .8 indicates a large effect.

```{r}
#find means for APA write-up/conclusion
mean(grades$Pretest)
sd(grades$Pretest)
mean(grades$Posttest)
sd(grades$Posttest)
```

APA conclusion: A related-samples t-test revealed a statistically significant difference between the pre-test (M=70.22, SD=9.22) and post-test (M=76.56, SD=9.30); t(8)=3.45, p<0.05, d=1.15. Cohen's d revealed a large effect.
