---
title: "Week 14 - Chi Square"
output: html_document
---

Week 14 Live session R code


This week we will learn how to conduct a chi-square test in R. We will start with the goodness-of-fit test (only ONE discrete variable). Let's look at the test format data from async:
```{r}
#create an object called "format" with three test values
format <- c(17,54,64)
```

We will run the chi-square test using the function `chisq.test`. We need to plug in the expected proportions for each group (all equally weighted in this example) - so we have 1/3 expected in each group.

```{r}
chisq.test(format, p=c(1/3, 1/3, 1/3))
#Note: unless you're given data for the expected frequencies, use an equal distribution of data
```
**APA conclusion**: A chi-square test of goodness-of-fit was performed to determine whether the three scoring methods were equally preferred. Preference for final scoring method was not equally distributed in the population, $\chi^2(2, N=135)=27.244$, $p<.05$.


Null: the distribution of colors of M&M candies are all the same
Alt: there are colors of M&M candies with different distributions

```{r}
MnMs <- c(18,12,11,17,16)
chisq.test(MnMs, p=c(1/5, 1/5, 1/5, 1/5, 1/5))
```
Conclusion: fail to reject the null. There is not sufficient evidence to show there are colors of M&M candies with significantly different distributions.

**APA conclusion**: A chi-square test of goodness-of-fit was performed to determine whether the colors of M&M candies are equally distributed. There was no evidence of a difference in colors, $\chi^2(4, N=74)=2.622$, $p>.05$.


Now, let's look at the test for independence (is there an association between two discrete variables).
```{r}
graduate <- read.table("graduate.txt", header=TRUE, sep="\t")

#label the levels of the variables for the table we're building (this is optional - but helpful to read your table)
graduate$Business <- factor(graduate$Business, levels= c("0","1"), labels = c("NonBiz", "Biz"))
graduate$Doctorate <- factor(graduate$Doctorate, levels= c("0","1"), labels = c("Masters", "Doctorate"))

#Now, let's summarize the data in a table (call it gradtable)
gradtable <- table(graduate$Business, graduate$Doctorate) #look at the table after you create it
gradtable
```

```{r}
chisq.test(gradtable, correct=F)
#Note: we use correct = F to get the same test statistic we computed by hand (no continuity correction)
```

Or we could input the values into the table.
```{r}
gradtable2 <- data.frame(c(574,176), c(185,3))

chisq.test(gradtable2, correct=F)
```

Based upon using a chi-square test for independence, we determined that there was an association between whether you have a business degree and whether you pursue a masters or a doctoral degree, X2(1)=46.57, p<0.05.


The Fisher Exact test is another nonparametric method for two discrete variables. In practice, it is often used instead of chi-square for tests of association/independence. The hypotheses are based on the odds ratios from the data contingency table.

Null: The odds ratio is equal to 1 (i.e. all groups are evenly divided) 
Alt: The odds ratio is not equal to 1

```{r}
#We'll use the same graduation data here
fisher.test(gradtable2)
```
As before, the p-value is less than .05, and so we reject the null hypothesis.


Null: there is no association between gender and voting party
Alt: there is an association between gender and voting party

```{r}
voters <- data.frame(c(200,150,50), c(250,300,50))

chisq.test(voters, correct = F)
```
Reject the null.
**APA Conclusion**: A chi-square test of independence was performed to examine the relation between gender and voting party. The relation between these variables was significant, $\chi^2(2, N=1,000)=16.204$, $p<.05$.


Now compare to the Fisher test:
```{r}
fisher.test(voters)
```
Again, we reach the same conclusion. The benefit of the Fisher's test is that it provides a more exact (or precise) estimate, compared to chi-square.

