---
title: "R Notebook"
output: html_document
---

# Week 12 Live session R code
This week we will learn how to conduct a SLR (simple linear regression) in R

Let's look at the PhD data from async:
```{r}
#setwd("C:/Users/blmos/Dropbox/School Things/VU Peabody/LLO 8180 - Applied Statistics/R Code")
phd <- read.table("phd.txt", header=TRUE, sep="\t")
```

Now we will make a scatterplot to examine the data visually
```{r}
plot(phd$Time, phd$Pubs, 
     main="Scatterplot of Time & Pubs", 
     xlab="Time Since Grad (in Years)", ylab="Publications")
```
Recall that we would only use linear regression for data that appears to have a linear relationship.


Let's run an SLR with Pubs as DV and Time as IV; we will save it to an object called phd.regression, and then view the summary results.
```{r}
phd.regression <- lm(Pubs ~ Time, data=phd)

summary(phd.regression)
```

This code will pull just the confidence interval for the coefficients
```{r}
#confidence intervals for parameter estimates
confint(phd.regression)
```

```{r}
#Plotting regression line on scatterplot
plot(phd$Time, phd$Pubs, 
     main="Scatterplot of Time & Pubs", 
     xlab="Time Since Grad (in Years)", ylab="Publications")
#add the regression line of best fit
abline(phd.regression, col="blue")
```
How well are our predictions? We can examine the residual (error) plot to see:
```{r}
phd.res<-resid(phd.regression)

plot(phd$Pubs, phd.res, main = "Residual Plot for Publications", xlab="Publications",ylab = "Error (Residual)")
abline(0,0, col="purple")
```

An additional example (using data that is stored in R)
I always look at the data first - dataset in R is called mtcars 

```{r}
summary(mtcars)
```

Let's examine a possible relationship between car weight & miles per gallon

```{r}
plot(mtcars$wt, mtcars$mpg, 
     main="Scatterplot of Weight and MPG", 
     xlab="Weight in 1000s", ylab="MPG")
```

Does car weight predict MPG? 

```{r}
cars.regression <- lm(mpg ~ wt, data=mtcars)
summary(cars.regression)

plot(mtcars$wt, mtcars$mpg, 
     main="Scatterplot of Weight and MPG", 
     xlab="Weight in 1000s", ylab="MPG")
#add the regression line of best fit
abline(cars.regression, col="red")
```
And here is the residual plot:
```{r}
cars.res<-resid(cars.regression)

plot(mtcars$mpg, cars.res, main = "Residual Plot for MPG", xlab="MPG",ylab = "Error (Residual)")
abline(0,0, col="purple")
```
