---
title: "Week 11 Correlation"
output: html_document
---

# Week 11 Live session R code

This week we will learn how to compute and analyze correlations


```{r}
#If you know your working directory, use the line below. Otherwise, use the menu: Session > Set Working Directory > Choose Directory
setwd("C:/Users/blmos/Dropbox/School Things/VU Peabody/LLO 8180 - Applied Statistics/R Code")
```

Let's look at the phd data from async. First, import the data:
```{r}
phd_data <- read.table("phd.txt", header=TRUE, sep="\t")

head(phd_data)
```
Note the variable names and spelling (remember capitalization matters!)


Now we will make a scatterplot:
```{r}
plot(phd_data$Time, phd_data$Pubs, main="Scatterplot of Time & Publications",
     xlab="Time Since PhD Grad (in years)", ylab="Publications", pch=19)
```

We can use R to compute the Pearson Correlation and obtain other values for the hypothesis test
```{r}
cor.test(phd_data$Time, phd_data$Pubs, method=c("pearson"))
```

Now, let's look the example from asyc for the Spearman correlation

```{r}
english_data <- read.table("english.txt", header=TRUE, sep="\t")

cor.test(english_data$Words, english_data$Grade, method=c("spearman"))
```

Point-Biserial Correlation (note that this code is the same as Pearson's, but the data setup differs)
```{r}
byod_data <- read.table("byod.txt", header=TRUE, sep="\t")

cor.test(byod_data$Grade, byod_data$Byod) #pt-biserial
```

Phi Correlation (note that this code is the same as Pearson's, but the data setup differs)
```{r}
grad_data <- read.table("graduate.txt", header=TRUE, sep="\t")

table(grad_data)

cor.test(grad_data$Doctorate, grad_data$Business) #phi
```
