---
title: "Week 4 - Normal Probabilities in Sampling Distributions"
author: "BL Mosby"
date: "6/3/21"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r  include=FALSE}
#This is our setup code chunk, where we would load any needed libraries, set document parameters, etc.
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This week we will learn how to compute probabilities of the sample mean, using the sampling distribution. 

Let's first use the R console as a calculator to compute total possible samples for theoretical sampling and then for experimental sampling.

How many samples of four participants (n = 4) from a population of ten people (N = 10) would we have using THEORETICAL sampling?

How many samples of four participants (n = 4) from a population of ten people (N = 10) would we have using EXPERIMENTAL sampling? (Use `n! = factorial(n)` for this formula) $$\frac{N!}{n!(N-n)!}$$

---
### Sampling Distribution of the Mean

Returning to the IQ score example: IQ tests are scored to have a population mean of 100 and standard deviation 15, and approximately normally distributed. Last week, we found the probability that an individual scores greater than 120 using two equivalent methods:
```{r}
#Method 1 - using the z-score in pnorm
(120-100)/15 #this finds the z-score
pnorm(1.33, lower.tail=FALSE)

#Method 2 - using the raw score in pnorm
pnorm(120, mean=100, sd=15, lower.tail=FALSE)
```

This week, we're going to work with the sampling distribution of the **mean**. In other words, instead of considering the probability value for an individual, we want to find the probability that the mean of a sample group is above or below a given value. 

**Central Limit Theorem** - the distribution of the sample mean is approximately normal, given the population is normally distributed OR the sample size is sufficiently large (i.e. > 30). Additionally,
$$\mu_M = \mu \text{ and} \\  \sigma_M \text{ or } SEM = \frac{s}{\sqrt{n}}$$
This enables us to compute the probability for the mean of a sample of individuals from a given population. For example, what is the probability that a sample of 9 individuals has a mean IQ greater than 110 on an IQ test?
 - First, we need to compute the standard error (se), as this will replace the standard deviation in our sampling distribution calculations. (Use the console to find se.)
 - Next, using either Method 1 or Method 2 above, find the probability using the CLT.

```{r}
#Type the code for the probability calculation here:

```

Now try the following questions out with different population values:

IQ tests are normed to have a mean of 140 and standard deviation of 12. What is the probability that a sample of 5 individuals has a mean less than 130 on an IQ test?

```{r}
#Type the code for the probability calculation here:

```

What's the probability that a sample of 10 has a mean greater than 150 on the same IQ test?
```{r}
#Type the code for the probability calculation here:

```

---
Now suppose we have state standardized math test, with mean of 170 and an sd of 15. 

What is the probability that a sample of 7 students has a mean greater than 175?
```{r}
#Type the code for the probability calculation here:

```
(What did we need to assume to complete this calculation?)


