---
title: "Applied Statistics - Assignment 5"
author: "YOUR NAME HERE"
date: "due 7/26/21"
output: 
  html_document: default
---

This is due prior to the beginning of live session on the due date. Please submit your knit file to the LMS. Round all reported statistics (when applicable) to the nearest hundredths place (i.e., two decimal places). 

Please identify students with whom you worked on this assignment here (MAX of four to a group):

1.[Week 9] To test whether arousal or stress levels increase as the difficulty of a task increases, eight participants were asked to complete an easy, typical or difficult task. Their galvanic skin response (GSR) was recorded. A GSR measures the electrical signals of the skin in units called microSiemens, with higher signals indicating greater arousal or stress. The data for each task is given in the dataset `difficulty.txt`.

a. Write the null hypothesis using a complete sentence, then as a mathematical statement.
b. Write the alternative hypothesis using a complete sentence.
c. Determine the degrees of freedom for within and between groups' variation. Show your calculations in the space below.

d. Run an ANOVA to obtain the test statistic and p-value.
```{r}

```

e. Make a decision about the null hypothesis.
f. If appropriate, calculate and interpret the effect size. If not, state why.
g. Conduct a post hoc analysis and interpret results, using complete sentences.
```{r}

```

h. Write the conclusion in APA style.


2.[Week 10] A developmental psychologist placed children in a social situation in which they were either rewarded or punished (Factor A: consequence) by a parent, sibling, or stranger (Factor B: type of adult). Following this social situation, children were placed back in the same social situation, and the time it took them (in seconds) to engage in the punished or rewarded behavior was recorded. The hypothetical data are given in the file `behavior.txt`.

a. Write the null hypotheses using complete sentences.
b. Write the alternative hypotheses using complete sentences.
c. Run an ANOVA to obtain the test statistics and p-values.
```{r}

```

d. Write complete sentence with a decision about *each* null hypothesis.
e. Write conclusion in APA style.
f. Create an interaction plot. Give a brief interpretation in context of the variables.
```{r}

```