---
title: "Applied Statistics - Assignment 7"
author: "YOUR NAME HERE"
date: "due 8/9/21"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This is due prior to the beginning of live session on the due date. Please submit your knit file (you can use html, pdf, or word) to the LMS. Round all reported statistics (when applicable) to the nearest hundredths place (i.e., two decimal places). 

Please identify students with whom you worked on this assignment here (MAX of four to a group):

## 1. [Week 13] The data (bloodpressure.txt) represent measures on a random sample of 25 individuals with high cholesterol levels. The variables are as follows: 
## Dependent variable Y: Systolic Blood Pressure (SBP)
## Independent variable X_1: Body Size, measured by Quetelet (QUET) Index 100(weight/height2)
## Independent variable X_2: Age

### a. Using R, carry out MLR analyses to obtain raw and standardized regression coefficients. Mean center age and QUET.
```{r}

```

#### i. Write the regression equation
#### ii. Interpret the intercept
#### iii. Interpret the slope for QUET
#### iv. Interpret the slope for age
#### v. Interpret the hypothesis tests for each regression coefficient
### b. Report the coefficient of multiple determination obtained in your MLR analysis and interpret the associated hypothesis test.
### c. For an individual of age 51 and QUET 3.30, obtain by hand calculation the predicted level of SBP (remember they should be mean centered). If that individual’s actual SBP was found to be 130, obtain the residual.
### d. Using R, obtain the squared semi-partial correlation of Age and SBP and interpret the value.
```{r}

```

### e. Using R, obtain the squared partial correlation of Age and SBP and interpret the value.
```{r}

```
